### 2012-04-02 doc/theses/mikula/proposed_algorithm.tex
Master thesis: Corrections.
Master thesis: In progress.

\section{Step 4: Key Discovery}
\todo[inline]{Tu by neskodilo napisat, ze sice preberame alg, ale ho aj implementujeme a overime, co v povodnom clanku nebolo.}
In this step, the algorithm discovers keys of elements, incorporating the approach from paper \cite{Necasky:2009:DXK:1529282.1529414}, described in Chapter \ref{Keys}, and extending it. Firstly, the syntax tree is searched for forms of FLWOR expressions to infer the keys from. Then, as described later, certain constructs are used to support or degrade the evidence of inferred statements in a final summary.

@@

\begin{algorithm}
\caption{Function \texttt{determineJoinPattern}}
\label{ALG_func_determineJoinPattern}
\begin{algorithmic}[1]
\REQUIRE{\ \\
    $curBindingNode$: A variable binding node in the current FLWOR node. \\
    $P$: A predicate from the binding expression. \\
    $forBindingNode$: A for binding node from an ancestor FLWOR node. \\
    $forVar$: A variable from the ancestor for binding node. \\
    $forVars$: An array of couples of for variables and their for clause nodes. \\
    $checkJoinPattern3$: A flag determining whether to analyze the join pattern as possible join pattern 3. \\
    $whereExpr$: An expression from a where clause of the current FLWOR node.
}

\IF{$P$ is of form $curBindingNode.varName/L_1 = var/L_2$}
    \IF{$is(curBindingNode,$ ForClauseNode$)$}
        \STATE memorize $(forBindingNode, curBindingNode)$ as an occurrence of the for join pattern
        \STATE remove $(forVar, forBindingNode)$ from $forVars$
    \ELSIF{$is(curBindingNode,$ LetClauseNode$)$}
        \STATE memorize $(forBindingNode, curBindingNode)$ as an occurrence of the for let pattern                
        \STATE remove $(forVar, forBindingNode)$ from $forVars$
    \ENDIF
\ELSIF{$checkJoinPattern3$ AND $is(curBindingNode,$ ForClauseNode$)$}
    \IF{$whereExpr$ is of form $bindingNode.varName/L_1 = var/L_2$}
        \STATE memorize $(forBindingNode, curBindingNode)$ as an occurrence of the join pattern 3
        \STATE remove $(forVar, forBindingNode)$ from $forVars$
    \ENDIF
\ENDIF

@@

When the algorithm discovers that a certain element is probably not unique, it assigns the element with a negative weight that will decrease the sum of its weights in the summarizing step.

@@

\IF{$is(node$, FunctionCallNode$)$}
    \IF{$node.fncName =$ 'distinct-values'}
        \STATE $argument := node.getChild(0)$
        \STATE $type := getType(argument)$
        \IF{$type$ is PathType \\ $\wedge$ $usesOnlyChildAndDescendantAndAttributeAxes(type)$ \\ $\wedge$ $isWithoutPredicatesExceptLastStep(type)$}
            \STATE memorize $argument$ as the negative key statement, weight 1
        \ENDIF
    \ENDIF
\ENDIF
\end{algorithmic}
\end{algorithm}

Besides the aggregation functions, a certain form of FLWOR expressions is utilized in this step. They are FLWORs that iterate through a sequence of nodes selected by a path expressions using only child and descendant axes, and either that path expressions end with a predicate comparing a value of a subnode with some other value (literal constant, reference of a variable which is constant in this expressions, etc), or, the comparison is located in a where clause of a respective FLWOR.

As discussed in Chapter \ref{Negative_uniqueness}, when a for clause is used to iterate through a set of items, it is expected that the number of the items may be more than one. And, together with that, the set is restricted to contain only items which satisfy a condition that a value of one of their subnodes (or descendant nodes) equals a certain value. That implies that the value in those nodes is not unique, and, alike the utilization of the aggregation functions, non-unique nodes cannot be keys.

@@

\IF{$is(exprNode,$ OperatorNode$)$}
    \IF{$exprNode.operator =$ GEN\_EQUALS}
        \FORALL{$var \in forVars$}
            \IF{$exprNode$ is of a form $var/L = C$ where $getType(var/L)$ is PathType $\wedge$ $isWithoutPredicatesExceptLastStep(getType(var/L))$ $\wedge$ $usesOnlyChildAndDescendantAndAttributeAxes(getType(var/L))$ $\wedge$ $C$ is a literal constant}
                \STATE memorize $var/L$ as the negative key statement, weight 0.9
            \ENDIF
        \ENDFOR
    \ELSIF{$exprNode.operator =$ AND}
        \STATE $processWhere(forVars, exprNode.leftSide)$
        \STATE $processWhere(forVars, exprNode.rightSide)$
    \ENDIF
\ENDIF
\end{algorithmic}
\end{algorithm}

\todo[inline]{Definicia NUS a context Path, podobne ako v klucoch.}

>>>>

\section{Step 4: Key Discovery}
\todo[inline]{Tu by neskodilo napisat, ze sice preberame alg, ale ho aj implementujeme a overime, co v povodnom clanku nebolo.}
In this step, the algorithm discovers keys of elements, incorporating the approach from paper \cite{Necasky:2009:DXK:1529282.1529414}, described in Chapter \ref{Keys}, and extending it. Despite the approach was proposed, it has not been implemented yet. Therefore we will be the first to implement it and perform consecutive testing.

Firstly, the syntax tree is searched for forms of FLWOR expressions to infer the keys from. Then, as described later, certain constructs are used to support or degrade the evidence of inferred statements in a final summary.

@@

\begin{algorithm}
\caption{Function \texttt{determineJoinPattern}}
\label{ALG_func_determineJoinPattern}
\begin{algorithmic}[1]
\REQUIRE{\ \\
    $curBindingNode$: A variable binding node in the current FLWOR node. \\
    $P$: A predicate from the binding expression. \\
    $forBindingNode$: A for binding node from an ancestor FLWOR node. \\
    $forVar$: A variable from the ancestor for binding node. \\
    $checkJoinPattern3$: A flag determining whether to analyze the join pattern as possible join pattern 3. \\
    $whereExpr$: An expression from a where clause of the current FLWOR node.
}

\IF{$P$ is of form $forVar/L_1 = curBindingNode.varName/L_2 \wedge usesOnlyChildAndDescendantAndAttributeAxes(L_1) \wedge usesOnlyChildAndDescendantAndAttributeAxes(L_2)$}
    \IF{$is(curBindingNode,$ ForClauseNode$)$}
        \STATE memorize $(forBindingNode, curBindingNode)$ as an occurrence of the for join pattern
    \ELSIF{$is(curBindingNode,$ LetClauseNode$)$}
        \STATE memorize $(forBindingNode, curBindingNode)$ as an occurrence of the for let pattern                
    \ENDIF
\ENDIF
\IF{$checkJoinPattern3 \wedge is(curBindingNode,$ ForClauseNode$) \wedge usesOnlyChildAndDescendantAndAttributeAxes(L_1) \wedge usesOnlyChildAndDescendantAndAttributeAxes(L_2)$}
    \IF{$whereExpr$ is of form $forVar/L_1 = bindingNode.varName/L_2$}
        \STATE memorize $(forBindingNode, curBindingNode)$ as an occurrence of the join pattern 3
    \ENDIF
\ENDIF
\end{algorithmic}
\end{algorithm}

@@

When the algorithm discovers that a certain element (or attribute) is probably not unique, it assigns the element (or attribute) with a negative weight that will decrease the sum of its weights in the summarizing step.

We can formally define negative uniqueness statements as follows:

\begin{define}[Negative uniqueness Statement]
A negative uniqueness statement is a construct $(C, P)$, where $C$ and $P$ are PathType instances that use only child, descendant, and descendant or self ($P$ also attribute) axes. $C$ is called \emph{context path}, $P$ \emph{target path}. $C$ can be omitted as in the case of keys.
\end{define}

Note that the restrictions placed on the paths $C$ and $P$ are less restrictive as in the case of keys. If we found a negative uniqueness statement ($C$, $P$), where $P$ or $C$ contains predicates, then it implies that also $(C', P')$ is a negative uniqueness statement, where $C'$ and $P'$ are $C$ and $P$ stripped of predicates. In the case of special function calls, the explanation is the same.

@@

\IF{$is(node$, FunctionCallNode$)$}
    \IF{$node.fncName =$ 'distinct-values'}
        \STATE $argument := node.getChild(0)$
        \STATE $type := getType(argument)$
        \IF{$type$ is PathType \\ $\wedge$ $usesOnlyChildAndDescendantAndAttributeAxes(type)$ \\ $\wedge$ $isWithoutPredicatesExceptLastStep(type)$}
            \STATE memorize $(argument)$ as the negative key statement, weight 1
        \ENDIF
    \ENDIF
\ENDIF
\end{algorithmic}
\end{algorithm}

Besides the aggregation functions, a certain form of FLWOR expressions is utilized in this step. They are FLWORs that iterate through a sequence of nodes selected by a path expressions using only child and descendant axes, and either that path expressions end with a predicate comparing a value of a subnode with some other value (literal constant, reference to a variable which is constant in this expressions, etc), or, the comparison is located in a where clause of a respective FLWOR.

As discussed in Chapter \ref{Negative_uniqueness}, when a for clause is used to iterate through a set of items, it is expected that the number of the items may be more than one. And, together with that, the set is restricted to contain only items which satisfy a condition that a value of one of their subnodes (or descendant nodes) equals a certain value. That implies that the value in those nodes is not unique, and, alike the utilization of the aggregation functions, non-unique nodes cannot be keys. This situation is handled in Algorithms \ref{ALG_rejection_of_uniqueness_comparison_with_a_constant} and \ref{ALG_rejection_of_uniqueness_function_process_where}.

@@

\IF{$is(exprNode,$ OperatorNode$)$}
    \IF{$exprNode.operator =$ GEN\_EQUALS}
        \FORALL{$var \in forVars$}
            \IF{$exprNode$ is of a form $var/L = C$ where $getType(var/L)$ is PathType $\wedge$ $isWithoutPredicatesExceptLastStep(getType(var/L))$ $\wedge$ $usesOnlyChildAndDescendantAndAttributeAxes(getType(var/L))$ $\wedge$ $C$ is a literal constant}
                \STATE memorize $(var/L)$ as the negative key statement, weight 0.9
            \ENDIF
        \ENDFOR
    \ELSIF{$exprNode.operator =$ AND}
        \STATE $processWhere(forVars, exprNode.leftSide)$
        \STATE $processWhere(forVars, exprNode.rightSide)$
    \ENDIF
\ENDIF
\end{algorithmic}
\end{algorithm}

In spite of the presented algorithms search for global negative uniqueness statements ($C$ path is omitted) only, we included the context path in the definition for possible future extensions.
